`r params$project_title`

### 2021-02-03 reports/rmarkdown/03_integrated_datasets.Rmd
updates clustering and samples used

```{r, echo = FALSE}
PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/" #Path to the data

@@

samples_to_use <- colnames(counts)[(colSums(counts == 0)<3)]
datasets_all_samples <- lapply( datasets_all_samples , function(x) { x [ , samples_to_use] } )
metadata <- metadata[ samples_to_use , ]
saveRDS(datasets_all_samples,"../../results/datasets_all_samples.RDS")
```

@@

```{r, echo = FALSE,fig.width=10,fig.height=9}
# all_microbiome <- rbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
#                         datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
#                         datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
# all_microbiome <- t( t(all_microbiome) / colSums(all_microbiome) )
all_microbiome <- all_microbiome + rnorm( length(all_microbiome) , mean = 0, sd = .00000001)
all_microbiome <- all_microbiome[ rowSums(all_microbiome>0)>=2 ,  ]
dim(all_microbiome)

zscore <- t(apply(all_microbiome,1,function(i){scale(i,T,T)}))
zscore[ zscore > 4 ] <- 4
zscore[ zscore < -4 ] <- -4
rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)
mypar(mar=c(2,15,2,2))
top_var <- apply(all_microbiome,1,var)
top_var_t <- sort(top_var,decreasing = T)[100]
is_top_var <- top_var > top_var_t 
barplot(sort(top_var[is_top_var],decreasing = T)[30:1],horiz = T,las=1)
#Compute sample-wise correlations
cors <- cor(all_microbiome[is_top_var,], method = "pearson")
adj <- (1-cors)/2
# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )

@@

#Graph construction  
k <- 15 # k for KNN
P <- 0.2# prunning threshold of SNN

#Compute KNN graph from correlation matrix
knn <- RANN::nn2(adj,k = k)
# knn <- RANN::nn2(t(zscore),k = k)
nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }

#Compute SNN graph from KNN
# jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))
# 
# jSNN[1:10,1:10]
# 
# # k=10
# # P=0.1
# snn <- dbscan::sNN(x = adj, k = k, jp = T, approx = TRUE)
# snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
# snnj <- matrix(0, nrow(snn$id), nrow(snn$id))
# 
# # for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$dist[idx,] }
# for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$shared[idx,]/k }
# 
# for(idx in 1:nrow(snn$id) ){  snnj[idx,snn$id[idx,]] <- snn$shared[idx,] }
# 
# snnm[ snnj <= ceiling(k*P) ] <- 0

#Prune SNN graph based on the Jaccard index
# pSNN <- jSNN
# pSNN <- pSNN / max(pSNN)

# hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
#      border=NA,col="grey60",las=1,
#      xaxs="i",yaxs="i",
#      main="SNN jaccard",xlim=c(0,1))
# abline(v=P,col="red",lty=2,xpd=F)
# 
# hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
#      border=NA,col="grey60",las=1,
#      xaxs="i",yaxs="i",
#      main="correlation coeficient",xlim=c(-1,1))
# abline(v=0,col="red",lty=1,xpd=F)

N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
P <- t(nn) %*% nn
B <- N - nn
C <- t(B) %*% B
Q <- nrow(nn)*N - C
J <- P/Q
J[is.nan(J)] <- 0
J[J < 0.2] <- 0
colnames(J) <- colnames(all_microbiome)
rownames(J) <- colnames(all_microbiome)
# image(J[,nrow(nn):1],col = colorRampPalette(c("grey95","black"))(90),axes=F )
# set.seed(1)
# g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
# plot( g , vertex.label.cex=0.000001 , vertex.color = "grey" , 
#       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
#       vertex.size=10 )

pSNN[ pSNN < P ] <- 0
pSNN[ pSNN > P ] <- cors[pSNN > P]
pSNN[ pSNN < 0 ] <- 0

g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
write.csv(J,"../../results/patient_SNN_graph.csv",row.names = T)

@@

set.seed(1)
# l <- layout_with_graphopt(g,niter = 1000)
set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=30)

cl <- igraph::cluster_louvain(g)
cl <- igraph::walktrap.community(g)
metadata$joint_clustering <- cl$membership
set.seed(1)
U <- uwot::umap(adj,n_neighbors = 30)
h <- hclust(as.dist(adj),method = "ward.D2")
# metadata$joint_clustering <- cutree(h,k = 5)
write.csv2(metadata,"../../results/metadata_integration.csv",row.names = T)

@@

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,main="umap",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
empty_plot()
legend("topleft",title = "Patient groups",
       legend = levels(factor(cl$membership)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M4_CVLv3)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,main="FR embedding",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M4_CVLv3)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,main="umap",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
empty_plot()
legend("topleft",title = "Patient groups\ngroup_M4_CVLv3",
       legend = levels(factor(metadata$group_M4_CVLv3)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
# title("Louvain clusters",cex.main=1,font.main=1)

@@

```{r, eval=F, echo = FALSE,fig.width=10,fig.height=4}

@@

```{r, eval=F, echo = FALSE,fig.width=10,fig.height=4}

datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) ,
                      rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) ,
                      rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))

groups <- factor( rep(metadata$joint_clustering,3) )
age <- rep(metadata$age,3)
work_time <- rep(metadata$sexwork_months,3)

datasets <- datasets[ colSums(all_microbiome)>0 ]
groups <- groups[ colSums(all_microbiome)>0 ]
age <- age[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

df <- data.frame(groups,datasets,age)

library(edgeR)
design <- model.matrix( ~ datasets * groups + age , data=df )
rownames(design) <- paste0(datasets,"_",colnames(all_microbiome))
colnames(all_microbiome) <- rownames(design)

y <- DGEList(counts = round(((2^all_microbiome)-1)*1000) , remove.zeros = T)
y <- calcNormFactors(y,method = "TMM")
y <- estimateGLMCommonDisp(y,design)
y <- estimateGLMTagwiseDisp(y,design)
fit <- glmFit(y, design)

#age
lrt <- glmLRT(fit,coef = 8)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
plot(top$logFC,-log10(top$PValue))

#age
lrt <- glmLRT(fit,coef = 4:7)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]

#age
lrt <- glmLRT(fit,coef = 9:12)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]

#age
lrt <- glmLRT(fit,coef = 13:16)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]

@@

cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=5,]))

@@

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
    edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
    vertex.size=10, main="graph layout FR",
    edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
    edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=mU,
    vertex.size=10, main="UMAP",
    edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
par(mar=c(0,0,0,0))
empty_plot()
legend("topleft",title = "bact. communities",
     legend = levels(factor(cl$membership)),
     bty = "n",pch = 21,pt.bg = pal,cex = .8,pt.cex = .9,xpd=T)

@@

```{r, echo = FALSE,fig.width  = 10, fig.height=12}
mypar(3,6,mar=c(1,1,2,1))
for(j in sort(unique(cl$membership)) ){
  empty_plot()
  text( 0,1,labels = paste0(rownames(cors)[ cl$membership == j ],collapse = "\n"),adj=c(0,1),cex=.7)
  title(main=paste0("community",j))
}
write.csv2(cl$membership,"../../results/bacterial_communities.csv",row.names = rownames(cors))

@@

# for( i in c("louvain_tissue_V3","louvain_CVL_V3","louvain_CVL_V2")){
#   for( j in c("ASV_tissue_V3_normalized_batch_corrected",
#           "ASV_CVL_V3_normalized_batch_corrected",
#           "ASV_CVL_V2_normalized_NOT_batch_corrected")){
#     temp_group <- as.character(df[,i])
#     temp_group[ is.na(temp_group) ] <- "NA"
#     temp_group[ temp_group == "" ] <- "NA"
#     temp_group <- factor(temp_group)
#     
#     barlist(data=module_means_per_dataset[[j]],
#             genes=rownames(module_means_per_dataset[[j]]),
#             labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
#             clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
#   }
# }

for( i in c("group_M4_CVLv3","group_M6_CVLv3","group_M9_CVLv3","joint_clustering")){
  for( j in c("ASV_tissue_V3_normalized_batch_corrected",
          "ASV_CVL_V3_normalized_batch_corrected",
          "ASV_CVL_V2_normalized_NOT_batch_corrected")){

@@

  TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
  bac_dataset <- module_means_per_dataset[[i]]
  dge <- read.csv2(paste0(PATH,"/results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
  samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i),] > 0)==2 ]

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/h.all.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgseaMultilevel(pathways = gmt,stats = sort(cors[,x],decreasing = T))
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- paste0( "bac.com.", bacteria_use )

  # pvalues[ pvalues < -log10(0.05)] <- 0
  # sort(rowSums(pvalues>0),T)
  # sort(colSums(pvalues>0),T)
  NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
  NESes <- t(as.data.frame(NESes))
  NESes[is.na(NESes)] <- 0
  NESes[ pvalues < -log10(0.05) ] <- 0
  # NESes <- NESes[rowSums(NESes!=0) >= 2 , colSums(NESes!=0) >= 2]

  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
  # o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
  # o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order
  # NESes <- NESes[o_bacs,o_kegg]

@@

  # image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
  #        axes=F,border=NA,main=i,xlab="",ylab="",line=.4,cex.main=1,font.main=1)
  # text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
  #       labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
  # text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
  #       labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)
  # add_letter(myletter); myletter <- myletter[-1]
  # empty_plot()
  # cat("\\clearpage")

@@

layout(matrix(c(1,2,3,4,4),nrow = 1))
par(mar=c(6,1,2,0))
for( i in names(NESse_list)){
  x <- NESse_list[[i]][,common_pathways]
  image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
         axes=F,border=NA,main=sub("_norm.*","",i),xlab="",ylab="",line=.4,cex.main=1,font.main=1)
  # add_letter(myletter); myletter <- myletter[-1]
  text(   seq(1,0,length.out = nrow(x)) , par("usr")[c(3)] ,
      labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
}
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)])
text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
      labels = sub("HALLMARK_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

@@

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
non_zeros <- colSums(TRX)!=0
TRX <- TRX [ top_vars_TRX , non_zeros ]
pathway_means <- t(sapply( colnames(x) , function(i){
  i <- gmt[[i]]
  i <- as.matrix( TRX[ rownames(TRX) %in% i , ] )
  if(ncol(i)>1){
    i <- ( i - apply(i,1,min) ) / ( apply(i,1,max)- apply(i,1,min) )
    i <- colMeans(i)
  } else {
    i <- (i - min(i)) / (max(i) - min(i))
  }
  return( i - min(i) )
} ))
plot_dots(data=pathway_means,main="patient groups",cex.main=1,font.main=1,
        genes=rownames(pathway_means),
        clustering=metadata$joint_clustering[non_zeros],
        srt=0,pal=c("blue","blue","navy","navy","grey95","grey95","firebrick3","black"))

par(mar=c(6,2,3,2))
image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
# add_letter(myletter); myletter <- myletter[-1]
text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
    labels = rev(rownames(x)), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
# empty_plot(xaxs="i",yaxs="i",
#            xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main=" ")
# text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
      # labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

@@

  TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
  bac_dataset <- module_means_per_dataset[[i]]
  dge <- read.csv2(paste0(PATH,"/results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
  samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i),] > 0)==2 ]

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgseaMultilevel(pathways = gmt,stats = sort(cors[,x],decreasing = T))
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- paste0( "bac.com.", bacteria_use )

  # pvalues[ pvalues < -log10(0.05)] <- 0
  # sort(rowSums(pvalues>0),T)
  # sort(colSums(pvalues>0),T)
  NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
  NESes <- t(as.data.frame(NESes))
  NESes[is.na(NESes)] <- 0
  NESes[ pvalues < -log10(0.05) ] <- 0
  # NESes <- NESes[rowSums(NESes!=0) >= 2 , colSums(NESes!=0) >= 2]

>>>>

```{r, echo = FALSE}
PATH <- "~/Desktop/Desktop_stuff/NBIS/SMS_Projects/broliden_5325/" #Path to the data

@@

samples_to_use <- colnames(counts)[counts[c(3),] != 0]
datasets_all_samples <- lapply( datasets_all_samples , function(x) { x [ , samples_to_use] } )
metadata <- metadata[ samples_to_use , ]

detected_counts <- detected_counts[,samples_to_use]
barlist( data = rbind( detected_counts),main = "detected features",
         genes = c(rownames(detected_counts)),labels = sub("_norm.*","",rownames(counts)),
         clustering = rep(1,nrow(metadata)),
         col="grey",draw_mean_lines=F)
add_letter("b")

saveRDS(datasets_all_samples,"../../results/datasets_all_samples.RDS")
```

@@

```{r, echo = FALSE,fig.width=10,fig.height=9}
# all_microbiome <- rbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
#                         datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
#                         datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
# all_microbiome <- t( t(all_microbiome) / colSums(all_microbiome) )
all_microbiome <- all_microbiome[ rowSums(all_microbiome>0) >= 3 ,  ]
# all_microbiome <- all_microbiome + rnorm( length(all_microbiome) , mean = 0, sd = .00000001)
dim(all_microbiome)

zscore <- t(apply(all_microbiome,1,function(i){scale(i,T,T)}))
zscore[ zscore > 4 ] <- 4
zscore[ zscore < -4 ] <- -4
rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)
mypar(mar=c(2,15,2,2))
top_var <- apply(all_microbiome,1,var)
top_var_t <- sort(top_var,decreasing = T)[100]
is_top_var <- top_var > top_var_t 
barplot(sort(top_var[is_top_var],decreasing = T)[30:1],horiz = T,las=1)

#Compute sample-wise correlations
cors <- cor(all_microbiome[,], method = "pearson")
adj <- (1-cors)/2
# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )

@@

#Graph construction  
k <- 10 # k for KNN
p <- 0.1 # prunning threshold of SNN

#Compute KNN graph from correlation matrix
knn <- RANN::nn2(adj,k = k)
nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }

#Computing SNN matrix
N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
P <- t(nn) %*% nn
B <- N - nn
C <- t(B) %*% B
Q <- nrow(nn)*N - C
J <- P/Q
J[is.nan(J)] <- 0
J[J < p] <- 0
colnames(J) <- colnames(all_microbiome)
rownames(J) <- colnames(all_microbiome)

#Transforming SNN into an igraph object
g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
write.csv(J,"../../results/patient_SNN_graph.csv",row.names = T)

@@

set.seed(1)
set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=30)

cl <- igraph::cluster_louvain(g)
# cl <- igraph::walktrap.community(g)

metadata$joint_clustering <- cl$membership
set.seed(1)
U <- uwot::umap(adj,n_neighbors = 30)
h <- hclust(as.dist(adj),method = "ward.D2")
# metadata$joint_clustering <- cutree(h,k = 5)
write.csv2(metadata,"../../results/metadata_integration.csv",row.names = T)

@@

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,main="umap",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
empty_plot()
legend("topleft",title = "Patient groups",
       legend = levels(factor(cl$membership)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M9_CVLv3)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,main="FR embedding",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M9_CVLv3)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,main="umap",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
empty_plot()
legend("topleft",title = "Patient groups\ngroup_M9_CVLv3",
       legend = levels(factor(metadata$group_M9_CVLv3)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
# title("Louvain clusters",cex.main=1,font.main=1)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M4_CVLv3)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,main="FR embedding",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M4_CVLv3)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,main="umap",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
empty_plot()
legend("topleft",title = "Patient groups\ngroup_M4_CVLv3",
       legend = levels(factor(metadata$group_M4_CVLv3)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
# title("Louvain clusters",cex.main=1,font.main=1)

@@

```{r}
mypar(mar=c(2,15,2,2))

barlist( data = rbind( detected_counts),main = "detected features",
         genes = c(rownames(detected_counts)),labels = sub("_norm.*","",rownames(counts)),
         clustering = metadata$joint_clustering,
         col=pal,draw_mean_lines=F)
add_letter("b")

barlist( data = all_microbiome, main = "detected features",
         genes = c("Gardnerella vaginalis",
                   "Lactobacillus iners",
                   "Lactobacillus jensenii",
                   "Lactobacillus crispatus/acidophilus",
                   "Prevotella amnii",
                   "Prevotella timonensis",
                   "Atopobium vaginae"),
         clustering = metadata$joint_clustering,
         col=pal,draw_mean_lines=F)

@@

@@

cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=10,]))

@@

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
    edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
    vertex.size=10, main="graph layout FR",
    edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
    edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=mU,
    vertex.size=10, main="UMAP",
    edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
par(mar=c(0,0,0,0))
empty_plot()
legend("topleft",title = "bact. communities",
     legend = levels(factor(cl$membership)),
     bty = "n",pch = 21,pt.bg = pal,cex = .8,pt.cex = .9,xpd=T)
```

@@

```{r, echo = FALSE,fig.width  = 10, fig.height=12}
mypar(3,6,mar=c(1,1,2,1))
for(j in sort(unique(cl$membership)) ){
  empty_plot()
  text( 0,1,labels = paste0(rownames(cors)[ cl$membership == j ],collapse = "\n"),adj=c(0,1),cex=.7)
  title(main=paste0("community",j))
}
write.csv2(cl$membership,"../../results/bacterial_communities.csv",row.names = rownames(cors))
```

@@

for( i in c("group_M4_CVLv3","group_M9_CVLv3","joint_clustering")){
  for( j in c("ASV_tissue_V3_normalized_batch_corrected",
          "ASV_CVL_V3_normalized_batch_corrected",
          "ASV_CVL_V2_normalized_NOT_batch_corrected")){

@@

  TRX <- as.matrix(datasets_all_samples[["Tissue_RNAseq_V3_normalized"]])
  bac_dataset <- module_means_per_dataset[[i]]
  samples_TRX_and_CVL3 <- colnames(bac_dataset)

  # mypar(4,3,mar=c(4,2,2,2))
  #filter TRX dataset
  TRX <- TRX [ , samples_TRX_and_CVL3 ]
  TRX <- TRX [ rowSums(TRX>0)>= 5 , ]

  #filter bac_dataset dataset
  bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
  dim(bac_dataset)

  cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
  dim(cors)

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/h.all.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgseaMultilevel(pathways = gmt,stats = cors[,x])
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- paste0( "bac.com.", bacteria_use )

  allpathways <- unique(unlist(lapply(enrichments,function(x){x[,1]})))
  allpvals <- setNames(rep(NA,length(allpathways)),allpathways)

  pvalues <- lapply(enrichments,function(x){
    temp <- allpvals
    temp[x$pathway] <- x$pval
    return(temp)
  })
  pvalues <- t(as.data.frame(pvalues))
  pvalues <- -log10( pvalues )
  pvalues[is.na(pvalues)] <- 1

  allNES <- setNames(rep(NA,length(allpathways)),allpathways)
  NESes <- lapply(enrichments,function(x) {
    temp <- allNES
    temp[x$pathway] <- x$NES
    return(temp)
  })
  NESes <- t(as.data.frame(NESes))
  NESes[is.na(NESes)] <- 0
  NESes[ pvalues < -log10(0.1) ] <- 0

@@

@@

layout(matrix(c(1,2,3,4,4),nrow = 1))
par(mar=c(6,1,2,0))
for( i in names(NESse_list)){
  x <- NESse_list[[i]][,common_pathways]
  image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
         axes=F,border=NA,main=sub("_norm.*","",i),xlab="",ylab="",line=.4,cex.main=1,font.main=1)
  # add_letter(myletter); myletter <- myletter[-1]
  text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
      labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
}
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)])
text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
      labels = sub("HALLMARK_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

@@

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
non_zeros <- colSums(TRX)!=0
TRX <- TRX [  , non_zeros ]
pathway_means <- t(sapply( colnames(x) , function(i){
  i <- gmt[[i]]
  i <- as.matrix( TRX[ rownames(TRX) %in% i , ] )
  if(ncol(i)>1){
    i <- ( i - apply(i,1,min) ) / ( apply(i,1,max)- apply(i,1,min) )
    i <- colMeans(i)
  } else {
    i <- (i - min(i)) / (max(i) - min(i))
  }
  return( i - min(i) )
} ))
plot_dots(data=pathway_means,main="patient groups",cex.main=1,font.main=1,
        genes=rownames(pathway_means),
        clustering=factor(metadata$joint_clustering[non_zeros]),
        srt=0,pal=c("blue","blue","navy","navy","grey95","grey95","firebrick3","black"))

@@

  TRX <- as.matrix(datasets_all_samples[["Tissue_RNAseq_V3_normalized"]])
  bac_dataset <- module_means_per_dataset[[i]]
  samples_TRX_and_CVL3 <- colnames(bac_dataset)

  # mypar(4,3,mar=c(4,2,2,2))
  #filter TRX dataset
  TRX <- TRX [ , samples_TRX_and_CVL3 ]
  TRX <- TRX [ rowSums(TRX>0)>= 5 , ]

  #filter bac_dataset dataset
  bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
  dim(bac_dataset)

  cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
  dim(cors)

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgseaMultilevel(pathways = gmt,stats = cors[,x])
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- paste0( "bac.com.", bacteria_use )

  allpathways <- unique(unlist(lapply(enrichments,function(x){x[,1]})))
  allpvals <- setNames(rep(NA,length(allpathways)),allpathways)

  pvalues <- lapply(enrichments,function(x){
    temp <- allpvals
    temp[x$pathway] <- x$pval
    return(temp)
  })
  pvalues <- t(as.data.frame(pvalues))
  pvalues <- -log10( pvalues )
  pvalues[is.na(pvalues)] <- 1

  allNES <- setNames(rep(NA,length(allpathways)),allpathways)
  NESes <- lapply(enrichments,function(x) {
    temp <- allNES
    temp[x$pathway] <- x$NES
    return(temp)
  })
  NESes <- t(as.data.frame(NESes))
  NESes[is.na(NESes)] <- 0
  NESes[ pvalues < -log10(0.1) ] <- 0
